---
title: "Data Science & Business Analytics Task_1"
author: "Nowful Ahamed"
output: html_document
---
##Liner Regression
  
  In this section we will see how the machine learning can be used to implement regression functions. We will start with simple linear regression involving two variables.

##Step 1:

Loading the necessary  library packages for this dataset. 
  readr     <- Use to read the .csv files.
  ggplot2   <- Use to create a plot. 
```{r}
library(readr)
library(ggplot2)
```

##Step 2:

Load the data set into R 
```{r}
Mark <- read.csv("student_mark.csv")
Mark
```

##Step 3:

Checking the summary of the variables
```{r}
summary(Mark)
```
Our variables are quantitative, when we run this function it tells us the minimum, median, mean, and maximum values of the independent variable (Score) and dependent variable (Hours):

##Step 4:

Make sure your data meet the assumptions for linear regression.The relationship between the independent and dependent variable must be    linear. We can test this visually with a scatter plot to see if the distribution of data points could be described with a straight line.
```{r}
data_view <-ggplot(Mark, aes(x=Hours,y=Score,col = Score))+ geom_point()
data_view
```

Now that you’ve determined your data meet the assumptions, you can perform a linear regression analysis to evaluate the relationship between the independent(Score) and dependent(Hours) variables.

##Step 5: 

For better prediction. Splitting the data as training and testing We are choosing more percentage of data  as training set and balance is testing data set   
```{r}
set.seed(0)
train_dummy <- sample(1:nrow(Mark),20)
train_data <- Mark[train_dummy,]
test_data <-Mark[-train_dummy,]
```
Now that data has splint into training(80%) and testing(20%),let start model building
```{r}
dim(train_data)
dim(test_data)
head(train_data)
head(test_data)
```

##Step 6:

Model building using linear regression  
```{r}
Model_1 <- lm(Score ~ Hours,data = train_data)
Model_1
```
##Step 7: 

  1) Predict the percentage of an student based on the number of study hours.
```{r}
predict_percentage <- predict(Model_1,newdata = test_data)
predict_percentage
plot(Mark$Hours,Mark$Score,main='HouseVsScore',col = 'blue')
abline(Model_1,col='green',lwd=2)
```
 
 2) What will be the predict score if a student study for 9.25 hrs/day.
```{r}
predict_score <- predict(Model_1,data.frame(Hours=9.25))
predict_score
```
The predicted Score: 92

##Step 8: 

  Now the final step to evaluate the performance of the model,For simplicity here, we have chosen the mean square error.
```{r}
mean((test_data$Score - predict_percentage)^2)
mean(abs(test_data$Score - predict_percentage))
```
Mean Absolute Error: 4.706122 

                                                                                                                       